---
title: "Regular Expressions and String Functions"
author: "Anil Akyildirim"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

## Question 1

Copy the introductory example. The vector name stores the extracted names.

R> name [1] "Moe Szyslak" "Burns, C. Montgomery" "Rev. Timothy Lovejoy" [4] "Ned Flanders" "Simpson, Homer" "Dr. Julius Hibbert"

(a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name. 

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.). 

(c) Construct a logical vector indicating whether a character has a second name. 

## Answer 1

```{r}
# installing required package 
install.packages('stringr', repos = "http://cran.us.r-project.org")
library('stringr')
```

```{r}
# creating the example raw data
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555 -6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
raw.data

```

```{r}
#creating the example name list
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name

```

**Answer (a)**

When we look at the name list we can see that we need to pay attention.

* Title of a person

* Spaces

* Order of the first name and the last name

* Middle Initial of an individual.

* Commas


```{r}
# removing the comma
names_without_comma <- str_replace_all(name, pattern = ",", replacement = "")
names_without_comma
```


```{r}
# removing the title
names_without_titles <- str_replace_all(names_without_comma, pattern = "[[:alpha:]]{2,}\\. ", replacement = "")
names_without_titles
```

```{r}
# removing the middle names
names_without_middle_names <- str_replace_all(names_without_titles, pattern = "[[:alpha:]]{1,}\\. ", replacement = "" )
names_without_middle_names


```

```{r}
# displaying the first_name _last_name
first_name_last_name <- str_replace(names_without_middle_names, pattern = "(\\w+)\\s+(\\w+)", replacement = "\\2 \\1")
first_name_last_name

```

**Answer (b) Logical Vector indicating whether a character has a title**

```{r}
#using str_detect to give us True or False of title existence
str_detect(names_without_comma, "[[:alpha:]]{2,}\\.")

```

**Answer (c) Logical Vector indicating whether a character has a middle name**

```{r}
# using str_detect to give us True of False of middle name existence 
str_detect(names_without_titles, "[[:alpha:]]{1,}\\.")

```


## Question 2

Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression. 

(a) [0-9]+\\$ 

(b) \\b[a-z]{1,4}\\b 

(c) .*?\\.txt$ 

(d) \\d{2}/\\d{2}/\\d{4} 

(e) <(.+?)>.+?</\\1>

## Answer 2

**Answer (a)**

This string that has one or more numbers between 0 and 9 and will end with $
so something like this  - 589$ -  


```{r}
web_text <- "In Europe they use $ sign at the end of the numbers, for example 435$"
web_text

```

```{r}

str_extract_all(web_text, "[0-9]+\\$")

```

**Answer(b)**

The string that has any character counts from 1 to 4. (words that has 1 to 4 letters/characters)

```{r}

str_extract_all(web_text, "\\b[a-z]{1,4}\\b")

```

**Answer(c)**

The string that ends with .txt

```{r}

file_formats <- "We have a lot of files named info.txt"
file_formats
```

```{r}

str_extract_all(file_formats, ".*?\\.txt$")

```


**Answer(d)**

The string consists of 2 digit number followed by "/", 2 digit number follwed by "/" and 4 digit number. Looks like date/month/year or month/date/year.

```{r}

web_date <- "The last day for enrollment to the class was 08/27/2019"

str_extract_all(web_date, "\\d{2}/\\d{2}/\\d{4}")


```

**Answer(e)**

The string consists of <...> and anything in between and ends with </> . Looks like html.

```{r}

web_code <- "The way you construct the html body is that you start with <p>These pretzels are making me thirsty!</p>"
str_extract_all(web_code, "<(.+?)>.+?</\\1>")

```


## Question 9

The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.

clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5 fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr

## Answer 9

```{r}
# creating the message object

scripted_message <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5 fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

scripted_message

```


```{r}
#lets see if we can find anything in the numbers.

str_extract_all(scripted_message, "[:digit:]")

```

```{r}

#lets see if we can find anything in the capital numbers.
str_extract_all(scripted_message, "[:upper:]")

```

Well it says Congratulations You are a super nerd but all characters are individually shown,

```{r}
# combining each separate character, adding period
cat(unlist(str_extract_all(scripted_message, "[[:upper:].]")))

```



